---
title: "Assessment of Project Data"
description: |
  Testing if data sets chosen for the group project are appropriate and working well!
author:
  - name: Teo Jun Peng
    url: https://teojp3-is415.netlify.app/
date: 2021-10-13
output:
  distill::distill_article:
      self_contained: false
      toc: true
      toc_depth: 2
      toc_float: true
---

```{r setup, include=FALSE, eval=TRUE, echo=TRUE, message=FALSE, error=FALSE, fig.retina=3}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center", tidy.opts=list(width.cutoff=80),tidy=TRUE)
```

# Setting up R Environment

As usual, loading in the relevant R packages that are helpful in the analysis process.

```{r results='hide'}
packages = c('readr','sp', 'sf', 'rgdal', 'spNetwork', 'tmap', 'FNN', 'spatstat')
  for (p in packages)
  {
    if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Data Sources

The following data will be used and tested:

- [Drinking Fountains (Melbourne, Australia)](https://data.melbourne.vic.gov.au/City-Council/Drinking-fountains/h4ih-tzqs)
- [Public Toilets (Melbourne, Australia)](https://data.melbourne.vic.gov.au/City-Council/Public-toilets/ru3z-44we)
- [Green Spaces (Melbourne, Australia)](https://data.melbourne.vic.gov.au/People/Landmarks-and-places-of-interest-including-schools/j5vt-ppat)
- [Sports Facilities (Melbourne, Australia)](https://data.melbourne.vic.gov.au/People/Landmarks-and-places-of-interest-including-schools/j5vt-ppat)
- [Melbourne's Municipal Boundary](https://data.melbourne.vic.gov.au/Property/Municipal-boundary/ck33-yh8z)


# Importing in Data

## Create owin object of Study Area's boundary

Our chosen area of study is the central city area of **Melbourne located in Victoria, Australia.** The area is also known as **The City of Melbourne**, as officially named by their local government authority. 

An owin object will be created just in case we need to utilize data sets that expands beyond our study area.

```{r eval=FALSE}
melbourne <- st_read(dsn = "data/Melbourne", layer = "MelbourneBoundary") %>%
  st_transform(crs = 28355)

melbourne_sp = as(as_Spatial(melbourne), "SpatialPolygons")

melbourne_owin = as(melbourne_sp, "owin")
```

## Importing in Study Factors

### Network Data Set

We decided to use the Cycling Network as the network data as we wanted to visualise the analysis from a cyclist's point of view. The cycling network data set is taken from the [official website of the municipality of Melbourne](https://www.melbourne.vic.gov.au/Pages/home.aspx).

```{r}
network <- readOGR(dsn="data/BicycleNetwork",
                   layer="BicycleNetwork",
                   verbose = FALSE)

network <- spTransform(network,CRS("+init=epsg:28355"))

```

## Study Factors

For this study that illustrates an example of Network-Constrained Point Patterns Analysis, we would like to uncover whether the **drinking fountains in Melbourne, Australia** are distributed randomly **along the cycling network** and if not, what are the factors that affect their locations.

### Main Data: Drinking Fountains

The drinking fountain data set is also taken from the official website of the municipality of Melbourne.

```{r}
fountain <- read_csv("data/Drinking_fountains.csv")

fountain <- st_as_sf(fountain, coords = c("lon", "lat"), crs = 4326) %>%
    st_transform(crs = 28355)
```

### Cross K Factors

We will be using *three factors* to test out their correlation and relationship with the location of drinking fountains. The three factors are: **1) Public Toilets, 2) Green Spaces, 3) Sport Facilities.**

Luckily, the various data sets are also available on the official website of the municipality of Melbourne, making it easier to conduct the study.

```{r}
toilet <- read_csv("data/Public_toilets.csv")

toilet <- st_as_sf(toilet, coords = c("lon", "lat"), crs = 4326) %>%
    st_transform(crs = 28355)

sports <- read_csv("data/SportsFacility.csv")

sports <- st_as_sf(sports, coords = c("lon", "lat"), crs = 4326) %>%
    st_transform(crs = 28355)

green <- read_csv("data/GreenSpaces.csv")

green <- st_as_sf(green, coords = c("lon", "lat"), crs = 4326) %>%
    st_transform(crs = 28355)
```

# Initial EDA and Mapping

The locations of the various factors are:

* Drinking Fountains (Blue)

* Public Toilets (Red)

* Green Spaces (Green)

* Sports Facilities (Orange)

```{r}
tmap_mode('view')
tm_shape(toilet) +
  tm_dots(col = 'red') +
tm_shape(fountain) +
  tm_dots(col = 'blue') +
tm_shape(network) +
  tm_lines() +
tm_shape(green) +
  tm_dots(col='green') +
tm_shape(sports) +
  tm_dots(col='orange')
```

So far so good! The data sets used looks quite promising, and we shall test out Kernel Density with them later on in the assessment :) .













