---
title: "Hands-on Exercise 7"
description: |
  Learning about how to computte Global and Local Measure of Spatial Autocorrelation  (GLSA) using spdep package.
author:
  - name: Teo Jun Peng
    url: https://teojp3-is415.netlify.app/
date: 09-26-2021
output:
  distill::distill_article:
      self_contained: false
      toc: true
      toc_depth: 2
      toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

In this hands-on exercise, we will learn how to compute Global and Local Measure of Spatial Autocorrelation (GLSA) by using **spdep** package. 

Some new knowledge learnt from this exercise would be:

* compute Global Spatial Autocorrelation (GSA) statistics by using appropriate functions of **spdep package**
  + plot Moran scatterplot
  + compute and plot spatial correlogram using appropriate function of spdep package
* compute Local Indicator of Spatial Association (LISA) statistics for detecting clusters and outliers by using appropriate functions of **spdep** package
* compute Getis-Ord’s Gi-statistics for detecting hot spot or/and cold spot area by using appropriate functions of **spdep** package; and
* to visualise the analysis output by using **tmap** package

# Getting Started

## The Analytical Question

In spatial policy, one of the main development objective of the local govenment and planners is to ensure equal distribution of development in the province. Our task in this study, hence, is to apply appropriate spatial statistical methods to discover if development are even distributed geographically. If the answer is **No*, our next question will then be “is there signs of spatial clustering?”. And, if the answer for this question is yes, then our next question will be “where are these clusters?”

In this case study, we are interested to examine the spatial pattern of a selected development indicator (i.e. GDP per capita) of Hunan Provice, People's Republic of China.([https://en.wikipedia.org/wiki/Hunan](https://en.wikipedia.org/wiki/Hunan))

## Study Area and Data

Two datasets will be used;

* Hunan province administrative boundary layer at county level. This is a geospatial data set in ESRI shapefile format.

* Hunan_2012.csv: This csv file contains selected Hunan’s local development indicators in 2012.


## Setting the Analytical Tools

```{r}
packages = c('sf', 'spdep', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

# Getting Data into R

In this section, we will importy a geospatial data set and its associated attribute table into R environment. The geospatial data is in ESRI shapefile format and the attribute table is in csv fomat.

## Import shapefile into R Environment

The code chunk below uses *st_read()* of **sf** package to import Hunan shapefile into R. The imported shapefile will be **simple features** Object of **sf**.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```
## Importing CSV file into R Environment

Next, we will import Hunan_2012.csv into R by using *read_csv()* of **readr** package. The output is in R data frame class.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```
## Perform Relational Join

The code chunk below will be used to update the attribute table of *hunan’s* SpatialPolygonsDataFrame with the attribute fields of *hunan2012* dataframe. This is performed by using *left_join()* of **dplyr** package.

```{r}
hunan <- left_join(hunan,hunan2012)
```
## Visualising Regional Development Indicator

Now, we are going to prepare a basemap and a choropleth map that shows the distribution of GDPPC 2012 by using *qtm()* of **tmap** package.

```{r fig.width=12, fig.height=8}
basemap <- tm_shape(hunan) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan, "GDPPC")
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```
# Global Spatial Autocorrelation

In this section, we will learn how to compute global spatial autocorrelation statistics and to perform spatial complete randomness test for global spatial autocorrelation.

## Computing Contiguity Spatial Weights

Before we can compute the global spatial autocorrelation statistics, we need to construct spatial weights of the study area. The spatial weights is used to define the neighbourhood relationships between the geographical units (i.e. county) in the study area.

In the code chunk below, *poly2nb()* of **spdep** package is used to compute contiguity weight matrices for the study area. This function builds a neighbours' list based on regions with contiguous boundaries. 

Specifying this "queen" argument is not necessary because the default is set to TRUE, meaning this function will return a list of first order neighbours using the Queen criteria. However, if you do not want to compute Queen spatial weights, you can set the argument to FALSE too, which will compute using Rook's criteria.

So specifically for this study, the code chunk below is used to compute **Queen contiguity weight matrix**.

```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```
The summary report above shows that there are 88 area units in Hunan and the most connected area unit has 11 neighbours. There are two area units with only one neighbour.

## Row-standardised weights matrix

Next, we need to assign weights to each neighboring polygon. 

In our case, each neighboring polygon will be assigned equal weight (style=“W”). This is accomplished by assigning the fraction 1/(#ofneighbors) to each neighboring county then summing the weighted income values. 

While this is the most intuitive way to summaries the neighbors’ values, it has one drawback in that polygons along the edges of the study area will base their lagged values on fewer polygons thus potentially over- or under-estimating the true nature of the spatial autocorrelation in the data. For this example, we’ll stick with the style=“W” option for simplicity’s sake but note that other more robust options are available, notably style=“B”.

```{r}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

The zero.policy=TRUE option allows for lists of non-neighbors. This should be **used with caution** since the user may not be aware of missing neighbors in their dataset. 

However, if a zero.policy of FALSE is used, it would return an error instead, helping users sieve out missing neighbors in their dataset when required.

## Global Spatial Autocorrelation: Moran’s I (Analytical Method)

In this section, we will perform Moran’s I statistics testing by using *moran.test()* of **spdep** or AKA the analytical method.

The null hypothesis we will be testing will be; 

* H0: Distribution of developments in Hunan are spatially independent

```{r}
moran.test(hunan$GDPPC, listw=rswm_q, zero.policy = TRUE, na.action=na.omit)
```
Question: What statistical conclusion can you draw from the output above?

* From the given output:
  + The p-value is 0.000001095, which is very small and statistically significant.
  + Hence, we will reject the null hypothesis and say that the developments are not spatially independent AKA there are strong signs of clustering. Since the Moran I Statistic is positive and approaching 1, we can infer that it reveals signs of cluster spatial pattern.
  
## Computing Monte Carlo Moran’s I 

The analytical approach to the Moran’s I analysis benefits from being fast. But it may be sensitive to irregularly distributed polygons. A safer approach to hypothesis testing is to run an MC simulation using the *moran.mc()* function. The moran.mc function takes an extra argument n, the number of simulations.

The code chunk below performs permutation test for Moran’s I statistic. A total of 1000 simulations will be performed.

*set.seed()* function sets the starting number used to generate a sequence of random numbers – it ensures that you get the same result if you start with that same seed each time you run the same process

```{r}
set.seed(1234)
bperm= moran.mc(hunan$GDPPC, listw=rswm_q, nsim=999, zero.policy = TRUE, na.action=na.omit)
bperm
```
Question: What statistical conclusion can you draw from the output above?

* Similar to the analytical method, since the p-value is small and statistically significant, we will reject the null hypothesis and say that the developments are not spatially independent AKA there are strong signs of clustering 

### Visualising Monte Carlo Moran’s I

It is always a good practice for us the examine the simulated Moran’s I test statistics in greater detail. This can be achieved by plotting the distribution of the statistical values as a histogram by using the code chunk below.

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, freq=TRUE, breaks=20, xlab="Simulated Moran's I")
abline(v=0, col="red") 
```
## Global Spatial Autocorrelation: Geary’s

We will now perform Geary’s c statistics testing by using appropriate functions of spdep package.

### Geary’s C test

The code chunk below performs Geary’s C test for spatial autocorrelation by using *geary.test()* of **spdep**.

```{r}
geary.test(hunan$GDPPC, listw=rswm_q)
```
Question: What statistical conclusion can you draw from the output above?

- Same as above conclusions, reject null hypothesis

### Computing Monte Carlo Geary’s C

```{r}
set.seed(1234)
bperm=geary.mc(hunan$GDPPC, listw=rswm_q, nsim=999)
bperm
```
Question: What statistical conclusion can you draw from the output above?

- Same as above conclusions, reject null hypothesis

### Visualising the Monte Carlo Geary’s C

Next, we will plot a histogram to reveal the distribution of the simulated values by using the code chunk below.

```{r}
mean(bperm$res[1:999])
```

```{r}
var(bperm$res[1:999])
```

```{r}
summary(bperm$res[1:999])
```

```{r}
hist(bperm$res, freq=TRUE, breaks=20, xlab="Simulated Geary c")
abline(v=1, col="red") 
```

## Spatial Correlogram

Spatial correlograms are great to examine patterns of spatial autocorrelation in your data or model residuals. They show how correlated are pairs of spatial observations when you increase the distance (lag) between them - they are plots of some index of autocorrelation (Moran’s I or Geary’s c) against distance.Although correlograms are not as fundamental as variograms (a keystone concept of geostatistics), they are very useful as an exploratory and descriptive tool. For this purpose they actually provide richer information than variograms.

### Compute Moran’s I correlogram

In the code chunk below, *sp.correlogram()* of spdep package is used to compute a 6-lag spatial correlogram of GDPPC. The global spatial autocorrelation used in Moran’s I. The **plot()** of base Graph is then used to plot the output.

```{r}
MI_corr <- sp.correlogram(wm_q, hunan$GDPPC, order=6, method="I", style="B")
plot(MI_corr)
```

### Compute Geary’s C correlogram and plot

In the code chunk below, *sp.correlogram()* of **spdep** package is used to compute a 6-lag spatial correlogram of GDPPC. The global spatial autocorrelation used in Geary’s C. The **plot()** of base Graph is then used to plot the output.

```{r}
GC_corr <- sp.correlogram(wm_q, hunan$GDPPC, order=6, method="C", style="W")
plot(GC_corr)
```

# Cluster and Outlier Analysis

Local Indicators of Spatial Association or LISA are statistics that evaluate the existence of clusters in the spatial arrangement of a given variable. For instance if we are studying cancer rates among census tracts in a given city local clusters in the rates mean that there are areas that have higher or lower rates than is to be expected by chance alone; that is, the values occurring are above or below those of a random distribution in space.

In this section, we will learn how to apply appropriate Local Indicators for Spatial Association (LISA), especially local Moran’I to detect cluster and/or outlier from GDP per capita 2012 of Hunan Province, PRC.

## Computing local Moran’s I

To compute local Moran’s I, the *localmoran()* function of **spdep** will be used. It computes *Ii* values, given a set of zi values and a listw object providing neighbour weighting information for the polygon associated with the zi values.

The code chunks below are used to compute local Moran’s I of GDPPC2012 at the county level.

```{r}
fips <- order(hunan$County)
localMI <- localmoran(hunan$GDPPC, rswm_q)
head(localMI)
```
*localmoran()* function returns a matrix of values whose columns are:

- Ii: the local Moran’s I statistics
- E.Ii: the expectation of local moran statistic under the randomisation hypothesis
- Var.Ii: the variance of local moran statistic under the randomisation hypothesis
- Z.Ii:the standard deviate of local moran statistic
- Pr(): the p-value of local moran statistic

The code chunk below list the content of the local Moran matrix derived by using *printCoefmat()*.

```{r}
printCoefmat(data.frame(localMI[fips,], row.names=hunan$County[fips]), check.names=FALSE)
```
## Mapping the local Moran’s I

Before mapping the local Moran’s I map, it is wise to append the local Moran’s I dataframe (i.e. localMI) onto hunan SpatialPolygonDataFrame. The code chunks below can be used to perform the task. The out SpatialPolygonDataFrame is called *hunan.localMI*.

```{r}
hunan.localMI <- cbind(hunan,localMI) %>%
  rename(Pr.Ii = Pr.z....E.Ii..)
```

### Mapping local Moran’s I values

Using choropleth mapping functions of **tmap** package, we can plot the local Moran’s I values by using the code chunks below.

```{r fig.width=12, fig.height=8}
tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty",
          palette = "RdBu",
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)
```

### Mapping local Moran’s I p-values

The choropleth shows there is evidence for both positive and negative Ii values. However, it is useful to consider the p-values for each of these values, as consider above.

The code chunks below produce a choropleth map of Moran’s I p-values by using functions of **tmap** package.

```{r fig.width=12, fig.height=8}
tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)
```

### Mapping both local Moran’s I values and p-values

For effective interpretation, it is better to plot both the local Moran’s I values map and its corresponding p-values map next to each other.

The code chunk below will be used to create such visualisation.

```{r fig.width=12, fig.height=8}
localMI.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Ii", 
          style = "pretty", 
          title = "local moran statistics") +
  tm_borders(alpha = 0.5)

pvalue.map <- tm_shape(hunan.localMI) +
  tm_fill(col = "Pr.Ii", 
          breaks=c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
          palette="-Blues", 
          title = "local Moran's I p-values") +
  tm_borders(alpha = 0.5)

tmap_arrange(localMI.map, pvalue.map, asp=1, ncol=2)
```

## Creating a LISA Cluster Map

The LISA Cluster Map shows the significant locations color coded by type of spatial autocorrelation. The first step before we can generate the LISA cluster map is to plot the Moran scatterplot.

### Plotting Moran scatterplot

The Moran scatterplot is an illustration of the relationship between the values of the chosen attribute at each location and the average value of the same attribute at neighboring locations.

The code chunk below plots the Moran scatterplot of GDPPC 2012 by using *moran.plot()* of **spdep**.

```{r}
nci <- moran.plot(hunan$GDPPC, rswm_q,
                  labels=as.character(hunan$County), 
                  xlab="GDPPC 2012", 
                  ylab="Spatially Lag GDPPC 2012")
```
Notice that the plot is split in 4 quadrants. The top right corner belongs to areas that have high GDPPC and are surrounded by other areas that have the average level of GDPPC. This are the high-high locations in the lesson slide.

### Plotting Moran scatterplot with standardised variable

First we will use *scale()* to centers and scales the variable. here centering is done by subtracting the mean (omitting NAs) the corresponding columns, and scaling is done by dividing the (centered) variable by their standard deviations.

```{r}
hunan$Z.GDPPC <- scale(hunan$GDPPC) %>% as.vector 
```

The *as.vector()* added to the end is to make sure that the data type we get out of this is a vector, that map neatly into out dataframe.

Now, we are ready to plot the Moran scatterplot again by using the code chunk below.

```{r}
nci2 <- moran.plot(hunan$Z.GDPPC, rswm_q,
                   labels=as.character(hunan$County),
                   xlab="z-GDPPC 2012", 
                   ylab="Spatially Lag z-GDPPC 2012")
```

### Preparing LISA map classes

The code chunks below show the steps to prepare a LISA cluster map.

```{r}
quadrant <- vector(mode="numeric",length=nrow(localMI))
```

Next, we centers the variable of interest around its mean.

```{r}
DV <- hunan$GDPPC - mean(hunan$GDPPC)     
```

This is follow by centering the local Moran’s around the mean.

```{r}
C_mI <- localMI[,1] - mean(localMI[,1])    
```

Next, we will set a statistical significance level for the local Moran.

```{r}
signif <- 0.05 
```

These four command lines define the high-high, low-low, low-high and high-low categories.

```{r}
quadrant[DV >0 & C_mI>0] <- 4      
quadrant[DV <0 & C_mI<0] <- 1      
quadrant[DV <0 & C_mI>0] <- 2
quadrant[DV >0 & C_mI<0] <- 3
```

Lastly, places non-significant Moran in the category 0.

```{r}
quadrant[localMI[,5]>signif] <- 0
```

In fact, we can combined all the steps into one single code chunk as shown below:

```{r}
quadrant <- vector(mode="numeric",length=nrow(localMI))
DV <- hunan$GDPPC - mean(hunan$GDPPC)     
C_mI <- localMI[,1] - mean(localMI[,1])    
signif <- 0.05       
quadrant[DV >0 & C_mI>0] <- 4      
quadrant[DV <0 & C_mI<0] <- 1      
quadrant[DV <0 & C_mI>0] <- 2
quadrant[DV >0 & C_mI<0] <- 3
quadrant[localMI[,5]>signif] <- 0
```

### Plotting LISA map

Now, we can build the LISA map by using the code chunks below.

```{r fig.width=12, fig.height=8}
hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)
```
For effective interpretation, it is better to plot both the local Moran’s I values map and its corresponding p-values map next to each other.

The code chunk below will be used to create such visualisation.

```{r fig.width=12, fig.height=8}
gdppc <- qtm(hunan, "GDPPC")

hunan.localMI$quadrant <- quadrant
colors <- c("#ffffff", "#2c7bb6", "#abd9e9", "#fdae61", "#d7191c")
clusters <- c("insignificant", "low-low", "low-high", "high-low", "high-high")

LISAmap <- tm_shape(hunan.localMI) +
  tm_fill(col = "quadrant", 
          style = "cat", 
          palette = colors[c(sort(unique(quadrant)))+1], 
          labels = clusters[c(sort(unique(quadrant)))+1],
          popup.vars = c("")) +
  tm_view(set.zoom.limits = c(11,17)) +
  tm_borders(alpha=0.5)

tmap_arrange(gdppc, LISAmap, asp=1, ncol=2)
```

# Hot Spot and Cold Spot Area Analysis

Beside detecting cluster and outliers, localised spatial statistics can be also used to detect hot spot and/or cold spot areas.

The term ‘hot spot’ has been used generically across disciplines to describe a region or value that is higher relative to its surroundings (Lepers et al 2005, Aben et al 2012, Isobe et al 2015).

## Getis and Ord’s G-Statistics

An alternative spatial statistics to detect spatial anomalies is the Getis and Ord’s G-statistics (Getis and Ord, 1972; Ord and Getis, 1995). 
It looks at neighbours within a defined proximity to identify where either high or low values cluster spatially. Here, statistically significant hot-spots are recognised as areas of high values where other areas within a neighbourhood range also share high values too.

*P.S. Variables used must only contain POSITIVE Values*

The analysis consists of three steps:

- Deriving spatial weight matrix
- Computing Gi statistics
- Mapping Gi statistics

## Deriving distance-based weight matrix

First, we need to define a new set of neighbours. Whist the spatial autocorrelation considered units which shared borders, for Getis-Ord we are defining neighbours based on distance.

There are two type of distance-based proximity matrix, they are:

- fixed distance weight matrix; and
- adaptive distance weight matrix.

### Deriving the centroid

We will need points to associate with each polygon before we can make our connectivity graph. It will be a little more complicated than just running *st_centroid()* on the sf object: **us.bound**. 

We need the coordinates in a separate data frame for this to work. To do this we will use a mapping function. The mapping function applies a given function to each element of a vector and returns a vector of the same length. 

Our input vector will be the geometry column of us.bound. Our function will be *st_centroid()*. We will be using map_dbl variation of map from the purrr package.

To get our longitude values we map the st_centroid() function over the geometry column of us.bound and access the longitude value through double bracket notation [[]] and 1. This allows us to get only the longitude, which is the first value in each centroid.

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
```

We do the same for latitude with one key difference. We access the second value per each centroid with [[2]].

```{r}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
```

Now that we have latitude and longitude, we use cbind to put longitude and latitude into the same object.

```{r}
coords <- cbind(longitude, latitude)
```

### Determine the cut-off distance

Firstly, we need to determine the upper limit for distance band by using the steps below:

- Return a matrix with the indices of points belonging to the set of the k nearest neighbours of each other by using *knearneigh()* of **spdep**.

- Convert the knn object returned by *knearneigh()* into a neighbours list of class nb with a list of integer vectors containing neighbour region number ids by using *knn2nb()*.

- Return the length of neighbour relationship edges by using *nbdists()* of **spdep**. The function returns in the units of the coordinates if the coordinates are projected, in km otherwise.

- Remove the list structure of the returned object by using **unlist()**.

```{r}
#coords <- coordinates(hunan)
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```
The summary output shows that the largest first nearest neighbour distance is 61.79 km, so using this as the upper threshold gives certainty that all units will have at least one neighbour.

### Computing fixed distance weight matrix

Now, we will compute the distance weight matrix by using *dnearneigh()* as shown in the code chunk below. 

- 62 is used after rounding up from the previous output of **Max Value 61.79**, so as to catch at least one neighbour per county

```{r}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

Next, *nb2listw()* is used to convert the nb object into spatial weights object. The input of nb2listw() must be an object of class ***nb.** The syntax of the function has two major arguments, namely style and zero.poly.

- *style* can take values “W”, “B”, “C”, “U”, “minmax” and “S”. B is the basic binary coding, W is row standardised (sums over all links to n), C is globally standardised (sums over all links to n), U is equal to C divided by the number of neighbours (sums over all links to unity), while S is the variance-stabilizing coding scheme proposed by Tiefelsdorf et al. 1999, p. 167-168 (sums over all links to n).

- If *zero policy* is set to TRUE, weights vectors of zero length are inserted for regions without neighbour in the neighbours list. These will in turn generate lag values of zero, equivalent to the sum of products of the zero row t(rep(0, length=length(neighbours))) %*% x, for arbitraty numerical vector x of length length(neighbours). The spatially lagged value of x for the zero-neighbour region will then be zero, which may (or may not) be a sensible choice.

```{r}
wm62_lw <- nb2listw(wm_d62, style = 'B')
summary(wm62_lw)
```

### Computing adaptive distance weight matrix

One of the characteristics of fixed distance weight matrix is that more densely settled areas (usually the urban areas) tend to have more neighbours and the less densely settled areas (usually the rural counties) tend to have lesser neighbours. Having many neighbours smoothes the neighbour relationship across more neighbours.

It is possible to control the numbers of neighbours directly using k-nearest neighbours, either accepting asymmetric neighbours or imposing symmetry as shown in the code chunk below.

```{r}
knn <- knn2nb(knearneigh(coords, k=8))
knn
```
Next, *nb2listw()* is used to convert the nb object into spatial weights object.

```{r}
knn_lw <- nb2listw(knn, style = 'B')
summary(knn_lw)
```

## Computing Gi statistics

### Gi statistics using fixed distance

```{r}
fips <- order(hunan$County)
gi.fixed <- localG(hunan$GDPPC, wm62_lw)
gi.fixed
```

The output of localG() is a vector of G or Gstar values, with attributes “gstari” set to TRUE or FALSE, “call” set to the function call, and class “localG”.

The Gi statistics is represented as a Z-score. Greater values represent a greater intensity of clustering and the direction (positive or negative) indicates high or low clusters.

Next, we will join the Gi values to their corresponding hunan sf data frame by using the code chunk below.

```{r}
hunan.gi <- cbind(hunan, as.matrix(gi.fixed)) %>%
  rename(gstat_fixed = as.matrix.gi.fixed.)
```

In fact, the code chunk above performs three tasks. First, it convert the output vector (i.e. *gi.fixed*) into r matrix object by using *as.matrix()*. Next, *cbind()* is used to join hun@data and *gi.fixed* matrix to produce a new SpatialPolygonDataFrame called **hunan.gi.* Lastly, the field name of the gi values is renamed to *gstat_fixed* by using *names()*.

### Mapping Gi values with fixed distance weights

The code chunk below shows the functions used to map the Gi values derived using FIXED distance weight matrix.

```{r fig.width=12,fig.height=8}
gdppc <- qtm(hunan, "GDPPC")

Gimap <-tm_shape(hunan.gi) +
  tm_fill(col = "gstat_fixed", 
          style = "pretty",
          palette="-RdBu",
          title = "local Gi") +
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, Gimap, asp=1, ncol=2)
```

### Gi statistics using adaptive distance

The code chunk below are used to compute the Gi values for GDPPC2012 by using an ADAPTIVE distance weight matrix (i.e *knb_lw*).

```{r}
fips <- order(hunan$County)
gi.adaptive <- localG(hunan$GDPPC, knn_lw)
hunan.gi <- cbind(hunan, as.matrix(gi.adaptive)) %>%
  rename(gstat_adaptive = as.matrix.gi.adaptive.)
```

### Mapping Gi values with adaptive distance weights

It is time for us to visualise the locations of hot spot and cold spot areas. The choropleth mapping functions of **tmap** package will be used to map the Gi values.

The code chunk below shows the functions used to map the Gi values derived using ADAPTIVE distance weight matrix.

```{r fig.width=12, fig.height=8}
gdppc <- qtm(hunan, "GDPPC")

Gimap <- tm_shape(hunan.gi) +
  tm_fill(col = "gstat_adaptive", 
          style = "pretty",
          palette="-RdBu",
          title = "local Gi") +
  tm_borders(alpha = 0.5)

tmap_arrange(gdppc, Gimap, asp=1, ncol=2)
```

Notice how the above output is much more smoother than its FIXED distance weight counterpart? Therefore, it really depends on context when choosing which method to use and the kind of visualisation one is looking for.








