---
title: "In-class Exercise 5"
description: |
  Learning more about Spatial Point Patterns Analysis.
author:
  - name: Teo Jun Peng
    url: https://teojp3-is415.netlify.app/
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE, eval=TRUE, echo=TRUE, message=FALSE, error=FALSE, fig.retina=3}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing and Loading the R package

- **maptools** is for creating spatial* objects
- **sf** is for importing geospatial data
- **raster** is for creating raster object
- **spatstat** is for performing SPPA
- **tmap** is plotting thematic maps
- **tidyverse** is for data handling

```{r}
packages <- c('maptools', 'sf', 'raster', 'spatstat', 'tmap', 'tidyverse', 'plotly', 'ggthemes')

for(p in packages){
  if(!require(p, character.only = TRUE)){
    install.packages(p)
  }
library(p, character.only = TRUE)
}
```

## Importing the Geospatial Data

Importing shapefile using *st_read*()* of **sf** package. The output object will be in tibble sf object class.

```{r}
mpsz_sf <- st_read(dsn = "data/ice5data/shapefile", 
                layer = "MP14_SUBZONE_WEB_PL")
```
Projection is in SVY21.

## Importing aspatial data from rds folder

*read_rds()* of **readr** package is used instead of readRDS() of base R is used. This is because output of *read_rds()* is in tibble object.

```{r}
childcare <- read_rds("data/ice5data/rds/childcare.rds")
chas <- read_rds("data/ice5data/rds/chas.rds")
```

Note that there are some data issues in childcare data frame because `Lat` and `Lng` should in numeric data type. The coordinate fields seem to be in decimal degrees. Hence, `WGS84` referencing system is assumed.

## Converting the aspatial data frame into sf objects

```{r}
CHAS_sf <- st_as_sf(chas,
                    coords = c("X_COORDINATE",
                               "Y_COORDINATE"),
                    crs = 3414)
```

Note: *st_as_sf()* accepts coordinates in character data types.

```{r}
childcare_sf <- st_as_sf(childcare,
                    coords = c("Lng",
                               "Lat"),
                    crs = 4326) %>%
  st_transform(crs = 3414)
```

## Plotting for reviewing

```{r}
tmap_mode("view")
tm_shape(childcare_sf) +
  tm_dots(alpha = 0.4,
          col = "blue",
          size = 0.05) +
tm_shape(CHAS_sf) +
  tm_dots(alpha = 0.4,
          col = "red",
          size = 0.05)
```

## Geospatial Data Wrangling

### Converting from sf to Spatial* classes

```{r}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
```

### Converting Spatial* data frame into Spatial* objects

*as.SpatialPoint()* and *as.SpatialPolygon()* of **maptools** package 

```{r}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
mpsz_sp <- as(mpsz, "SpatialPolygons")
```

### Converting from Spatial* objects into ppp objects

Using *as.ppp()* of **maptools** package

```{r}
childcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, 'ppp')
```

### Removing duplicate points using jitter

After using jitter, we then check for any duplicates left using the duplicated function.

```{r}
childcare_ppp_jit <- rjitter(childcare_ppp,
                             retry = TRUE,
                             nsim = 1,
                             drop = TRUE)
any(duplicated(childcare_ppp_jit))
```
```{r}
CHAS_ppp_jit <- rjitter(CHAS_ppp,
                             retry = TRUE,
                             nsim = 1,
                             drop = TRUE)
any(duplicated(CHAS_ppp_jit))
```
### Extracting Punggol Planning Area

Only extracting Punggol from mpsz, since that's the only area we want to study

Note: Need to put comma behind for function to run, if not will have syntax error. 

```{r}
pg <- mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL", ]
```

### Converting SpatialPolygonsDataFrame into SpatialPolygons object

```{r}
pg_sp <- as(pg, "SpatialPolygons")
```

### Converting SpatialPolygons into own object

```{r}
pg_owin <- as(pg_sp, "owin")
```

### Extracting spatial points within owin

```{r}
childcare_pg <- childcare_ppp_jit[pg_owin]
CHAS_pg <- CHAS_ppp_jit[pg_owin]
```

```{r}
plot(childcare_pg)
```

## L-function

```{r}
L_childcare <- envelope(childcare_pg,
                        Lest,
                        nsim = 99,
                        rank = 1,
                        gloval = TRUE)
```
```{r}
L_CHAS <- envelope(CHAS_pg,
                   Lest,
                   nsim = 99,
                   rank = 1,
                   gloval = TRUE)
```

# Code chunk for plotting interactive L-function

L_childcare

```{r}
title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_childcare)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
  }
```

L_CHAS

```{r}
title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_CHAS)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
  }
```










